### 2016-04-23 ch03-lrm/lrm-paper.tex
Stuff :-)

\citep{Pintelon2010LPM1},\citep{Pintelon2010LPM2},\citep{McKelvey2012LRM}

@@

\paragraph*{Outline}
\secref{sec:theory} introduces theory of estimating \glspl{FRF} and the relevant local modeling techniques.
In \secref{sec:simulations} the performance of the different modeling methods is investigated by means of simulations.
The methods are then illustrated on measurements of mechanical systems in \secref{sec:measurements}.
Finally, conclusions are presented in \secref{sec:conclusion}.

\begin{itemize}
    \item \TODO{choice $s$ or $z$}
    \item \TODO{influence of noise filter: mismatch with model}
\end{itemize}

@@

\TODO{hier voortschrijven}

@@

Do note that the estimated parameters $\theta \in \ComplexMatrix{\nth \times 1}$ but that $\ELSCost{\theta}$ is a real-valued function.
In contrast to a complex-valued function, this requires additional attention when the derivatives are calculated~\citep{Messerschmitt2006} and~\citep[Section 15.9]{Pintelon2012} since $\ELSCost{\theta}$ is not an analytic function.
Particularly, the function $\ELSCost{\theta}$ is regarded as a two-dimensional function $\ELSCost{\theta,\conj{\theta}}$ where $\conj{\theta}$ and $\theta$ are considered as independent variables~\citep{Hjorungnes2007,Hjorungnes2011}.
The derivatives are then computed with respect to these variables.
% Let us introduce the notations
% \begin{equation}
%   \ReIm{X} 
%   \isdef 
%   \begin{bmatrix}
%     \real{X}\\
%     \imag{X}
%   \end{bmatrix}
%   \quad  \text{,} \quad
%   \ReImImRe{X} 
%   \isdef 
%   \begin{bmatrix}
%     \real{X} & -\imag{X}\\
%     \imag{X} & \hphantom{-} \real{X}
%   \end{bmatrix}
% \end{equation}
% and the shorthand $\ReIm{\theta} = \ReImTheta$.
% E.g. in \citet[Section 15.9]{Pintelon2012} and \citet[Section 5.1]{Seber2007}, it is stated that $\ReImImRe{X}$ is isomorphic with $X$ such that operations on a complex matrix $X$ can be written as equivalent operations on the real matrix $\ReImImRe{X}$.
% As a result,
% \begin{align}
%   \Delta\theta &=
%   \left( \mat{1 & j} \kron \Identity{\nth[]}  \right) 
%   \ReIm{\Delta\theta}
%   \\
%   \ReIm{\Delta\theta} &= 
%   - \ReImImRe{ \left(   \PartialDerivative[2]{\ELSCost{\true\theta}}{\true\theta^2} \right)^{-1} }
%     \ReIm{ \left( \PartialDerivative{\ELSCost[n]{\true\theta}}{\true\theta} \right)^{\T}}
% \end{align}
% where $\kron$ denotes the Kronecker product~\citep{Brewer1978}.
To facilitate the derivations, we rewrite the expected value of the cost function in terms of $\theta$ and $\conj{\theta}$:
\begin{align}
   \ELSCost[0]{\theta,\conj{\theta}} & =
      \frac{1}{\nWind}
           \sum_{\dW}
           \LocalModel{Q}(      \dW,        \LocalVector{\theta},  \true{      \LocalVector{Y}},  \true{      \LocalVector{U} })
           \LocalModel{Q}(\conj{\dW}, \conj{\LocalVector{\theta}}, \true{\conj{\LocalVector{Y}}}, \true{\conj{\LocalVector{U}}})
   \\
   \LocalModel{Q}(\dW, \LocalVector{\theta},  \true{\LocalVector{Y}},  \true{ \LocalVector{U} }) & \isdef
                 \LocalModel{A}(\dW, \LocalVector{\theta}) \true{\LocalVector{Y}}(\dW)
               - \LocalModel{B}(\dW, \LocalVector{\theta}) \true{\LocalVector{U}}(\dW)
               - \LocalModel{C}(\dW, \LocalVector{\theta})
    \\             
    \ELSCost[n]{\theta,\conj{\theta}} & =
              \frac{\sigma_V^2}{\nWind} 
                     \sum_{r} 
                               \LocalModel{A}(      \dW,        \LocalVector{\theta} ) 
                               \LocalModel{A}(\conj{\dW}, \conj{\LocalVector{\theta}}) 
\end{align}
since $\conj{ \LocalModel{A}(\dW,\theta)} =  \LocalModel{A}(\conj{\dW},\conj{\theta}) $, and similar for $\LocalModel{B}$ and $\LocalModel{C}$ and $\LocalModel{Q}$.
Remember also that $\theta = \mat{\theta_A^{\TT} & \theta_B^{\TT} & \theta_C^{\TT}}$.

\paragraph{Contributions to the first order derivative}
The last term in \eqref{eq:nparam:bias-basic}, i.e. the first order derivative,  is given by
\begin{equation}
  \PartialDerivative{\ELSCost[0]{\true\theta, \conj{\true\theta}}}{\theta}
  = 
  \begin{bmatrix}
  \PartialDerivative{\ELSCost[0]{\true\theta, \conj{\true\theta}}}{\theta_A} &
  \PartialDerivative{\ELSCost[0]{\true\theta, \conj{\true\theta}}}{\theta_B} &
  \PartialDerivative{\ELSCost[0]{\true\theta, \conj{\true\theta}}}{\theta_C}
  \end{bmatrix}
\end{equation}
and split into the parts pertaining to respectively $\LocalModel{A}$, $\LocalModel{B}$ and, $\LocalModel{C}$.
The respective components are given by:
\begin{align}
  \PartialDerivative{\ELSCost[n]{\theta,\conj{\theta}}}{a_i} 
     &= 
     \frac{\sigma_V^2}{\nWind}
     \sum_{\dW} \dW^i \LocalModel{A}(\conj{\dW},\conj{\LocalVector{\theta}}) 
     && \forall i \in \set{1,\ldots,\order{A}}
     \\
   \PartialDerivative{\ELSCost[n]{\theta,\conj{\theta}}}{b_i} 
   &= 0
   && \forall i \in \set{0,\ldots,\order{B}}
   \\
   \PartialDerivative{\ELSCost[n]{\theta,\conj{\theta}}}{c_i} 
   &= 0
   &&\forall i \in \set{0,\ldots,\order{C}}
   \text{.}
\end{align}
It can be seen that only the factors pertaining to $\theta_{A}$ are present.
As this term occurs linearly in expression \eqref{eq:nparam:bias-basic}, only $\theta_A$ can incur a bias.
Consequently, only the poles can shifted by the presence of noise.
As such, the bias term $\Delta \theta$ can be written in the following sparse form:
\begin{equation}
  \Delta \theta = 
  \begin{bmatrix}
    \Delta \theta_A\\
    \Delta \theta_B\\
    \Delta \theta_C
  \end{bmatrix}
  =
  \begin{bmatrix}
    \Delta \theta_A\\
    \deemph{\Zero{\order{B}+1\times1}}\\
    \deemph{\Zero{\order{C}+1\times1}}
  \end{bmatrix}
  \text{.}
  \label{eq:nparam:lrm:bias:sparse}
\end{equation}

@@

However, the sparsity of $\Delta\theta$, see \eqref{eq:nparam:lrm:bias:sparse}, can be exploited such that the bias on the \gls{FRF} can be expressed as:
\begin{equation}
  \Delta \LocalModel{G}(\theta,\dW)
  \approx
  \frac{\LocalModel{G}(\theta, \dW)}
           {\LocalModel{A}(\theta, \dW)}
  \begin{bmatrix}
    \dW^1 & \deemph{\cdots} & \dW^{\order{A}}
  \end{bmatrix}
  \Delta \theta_A
  \text{.}
\end{equation}
Equivalently, the relative bias of the \gls{FRF} can be written as
\begin{equation}
\frac{\Delta \LocalModel{G}(\theta,\dw{0}{k})}{\LocalModel{G}(\theta,\dw{0}{k})}
\approx
\LocalModel{A}(\theta,\dW)^{-1}
\begin{bmatrix}
    \dw{0}{k}^1 & \deemph{\cdots} & \dw{0}{k}^{\order{A}}
  \end{bmatrix}
  H^{-1}
  \begin{bmatrix}

  \end{bmatrix}
\end{equation}

@@

\begin{table}
\centering
\caption{Local modeling methods used in the Monte Carlo analysis.}
\label{tbl:nparam:methods}
\begin{tabular}{lccccr} 
\toprule
\textsc{Method} & \textsc{Color} & $\nWind$ & $\nth$ & \DOF & $\Delta\SNR \axisunit{dB}$ \\
\midrule
 $\ETFE$ &\color{etfe} --- & 1  & 1  & 0 & 0 \\
 $\lrm{7,1,1,1}$ & \color{lrm7111} --- & 15  & 5 & 10  & 4.77 \\
 $\lric{7,2,2,2}$ &\color{lric7222} --- &   15 & 8 & 7  & 2.73 \\
 $\lrm{7,2,2,2}$ &\color{lrm7222} --- &   15 & 8 & 7  & 2.73 \\
 $\lrm{6,2,2,2}$ &\color{lrm6222} --- &   13 & 8 & 5  & 2.11 \\
 $\lpm{7,4,4}$ &\color{lpm744} --- &   15 & 10 & 5  & 1.76 \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{guideline}[Use the LRM when $\SNR \geq 20 \unit{dB}$ for resonant systems]
\end{guideline}

\begin{guideline}[Use the LPM when $\SNR \leq 20 \unit{dB}$]
\end{guideline}

\begin{guideline}[For long measurements, LPM and LRM are similar]
\end{guideline}

@@

  \section{Measurements}
  \label{sec:measurements}
   \subsection{AVIS}
   % \subsection{F16}

>>>>

@@

\paragraph*{Outline}
\secref{sec:theory} introduces theory of estimating \glspl{FRF} and the relevant local modeling techniques.
In \secref{sec:biascalc}, the bias of the \gls{LRM} is derived.
In \secref{sec:simulations} the performance of the different modeling methods are compared by means of simulations.
Finally, conclusions are presented in \secref{sec:conclusion}.
%TODO: choice s vs z
%TODO: influence of resonant noise filter

@@

\TODO{hier voortschrijven: ruisvariantie berekenen}

\begin{remark}
This system can only be solved reliably if the columns in \eqref{eq:nparam:designMatrix} are linearly independent.
In practice, this corresponds to the requirements on the input spectrum as stated in \citet{Schoukens2009LPM}: the input spectrum should be sufficiently `rough' to separate the transient contribution.
This can be obtained e.g. using random phase multisines or random input signals.
\end{remark}

@@

Do note that the estimated parameters $\theta \in \ComplexMatrix{\nth \times 1}$ but that $\ELSCost{\theta}$ is a real-valued function.
In contrast to a complex-valued function, this requires additional attention when the derivatives are calculated~\citep{Messerschmitt2006} and~\citep[Section 15.9]{Pintelon2012} since $\ELSCost{\theta}$ is not an analytic function.
Particularly, the function $\ELSCost{\theta}$ is regarded as a two-dimensional function $\ELSCost{\theta,\conj{\theta}}$ where $\conj{\theta}$ and $\theta$ are considered as independent variables~\citep{Hjorungnes2007,Hjorungnes2011}.
The derivatives are then computed with respect to these variables.
% Let us introduce the notations
% \begin{equation}
%   \ReIm{X} 
%   \isdef 
%   \begin{bmatrix}
%     \real{X}\\
%     \imag{X}
%   \end{bmatrix}
%   \quad  \text{,} \quad
%   \ReImImRe{X} 
%   \isdef 
%   \begin{bmatrix}
%     \real{X} & -\imag{X}\\
%     \imag{X} & \hphantom{-} \real{X}
%   \end{bmatrix}
% \end{equation}
% and the shorthand $\ReIm{\theta} = \ReImTheta$.
% E.g. in \citet[Section 15.9]{Pintelon2012} and \citet[Section 5.1]{Seber2007}, it is stated that $\ReImImRe{X}$ is isomorphic with $X$ such that operations on a complex matrix $X$ can be written as equivalent operations on the real matrix $\ReImImRe{X}$.
% As a result,
% \begin{align}
%   \Delta\theta &=
%   \left( \mat{1 & j} \kron \Identity{\nth[]}  \right) 
%   \ReIm{\Delta\theta}
%   \\
%   \ReIm{\Delta\theta} &= 
%   - \ReImImRe{ \left(   \PartialDerivative[2]{\ELSCost{\true\theta}}{\true\theta^2} \right)^{-1} }
%     \ReIm{ \left( \PartialDerivative{\ELSCost[n]{\true\theta}}{\true\theta} \right)^{\T}}
% \end{align}
% where $\kron$ denotes the Kronecker product~\citep{Brewer1978}.
To facilitate the derivations, we rewrite the expected value of the cost function in terms of $\theta$ and $\conj{\theta}$:
\begin{align}
   \ELSCost[0]{\theta,\conj{\theta}} & =
      \frac{1}{\nWind}
           \sum_{\dW}
           \LocalModel{Q}(      \dW,        \LocalVector{\theta},  \true{      \LocalVector{Y}},  \true{      \LocalVector{U} })
           \LocalModel{Q}(\conj{\dW}, \conj{\LocalVector{\theta}}, \true{\conj{\LocalVector{Y}}}, \true{\conj{\LocalVector{U}}})
   \\
   \LocalModel{Q}(\dW, \LocalVector{\theta},  \true{\LocalVector{Y}},  \true{ \LocalVector{U} }) & \isdef
                 \LocalModel{A}(\dW, \LocalVector{\theta}) \true{\LocalVector{Y}}(\dW)
               - \LocalModel{B}(\dW, \LocalVector{\theta}) \true{\LocalVector{U}}(\dW)
               - \LocalModel{C}(\dW, \LocalVector{\theta})
    \\             
    \ELSCost[n]{\theta,\conj{\theta}} & =
              \frac{\sigma_V^2}{\nWind} 
                     \sum_{r} 
                               \LocalModel{A}(      \dW,        \LocalVector{\theta} ) 
                               \LocalModel{A}(\conj{\dW}, \conj{\LocalVector{\theta}}) 
\end{align}
since $\conj{ \LocalModel{A}(\dW,\theta)} =  \LocalModel{A}(\conj{\dW},\conj{\theta}) $, and similar for $\LocalModel{B}$ and $\LocalModel{C}$ and $\LocalModel{Q}$.

\paragraph{Contributions to the first order derivative}
The last term in \eqref{eq:nparam:bias-basic}, i.e. the first order derivative,  is given by
\begin{equation}
  \PartialDerivative{\ELSCost[0]{\true\theta, \conj{\true\theta}}}{\theta}
  = 
  \begin{bmatrix}
  \PartialDerivative{\ELSCost[0]{\true\theta, \conj{\true\theta}}}{\theta_A} &
  \PartialDerivative{\ELSCost[0]{\true\theta, \conj{\true\theta}}}{\theta_B} &
  \PartialDerivative{\ELSCost[0]{\true\theta, \conj{\true\theta}}}{\theta_C}
  \end{bmatrix}
\end{equation}
and split into the parts pertaining to respectively $\LocalModel{A}$, $\LocalModel{B}$ and, $\LocalModel{C}$.
The respective components are given by:
\begin{align}
  \PartialDerivative{\ELSCost[n]{\theta,\conj{\theta}}}{a_i} 
     &= 
     \frac{\sigma_V^2}{\nWind}
     \sum_{\dW} \dW^i \LocalModel{A}(\conj{\dW},\conj{\LocalVector{\theta}}) 
     && \forall i \in \set{1,\ldots,\order{A}}
     \\
   \PartialDerivative{\ELSCost[n]{\theta,\conj{\theta}}}{b_i} 
   &= 0
   && \forall i \in \set{0,\ldots,\order{B}}
   \\
   \PartialDerivative{\ELSCost[n]{\theta,\conj{\theta}}}{c_i} 
   &= 0
   &&\forall i \in \set{0,\ldots,\order{C}}
   \text{.}
\end{align}
It can be seen that only the factors pertaining to $\theta_{A}$ are present.
As this term occurs linearly in expression \eqref{eq:nparam:bias-basic}, only $\theta_A$ can incur a bias.
Consequently, only the poles can shifted by the presence of noise.
As such, the bias term $\Delta \theta$ can be written in the following sparse form:
\begin{equation}
  \Delta \theta = 
  \begin{bmatrix}
    \Delta \theta_A\\
    \Delta \theta_B\\
    \Delta \theta_C
  \end{bmatrix}
  =
  \begin{bmatrix}
    \Delta \theta_A\\
    \deemph{0}\\
    \deemph{0}
  \end{bmatrix}
  \text{.}
  \label{eq:nparam:lrm:bias:sparse}
\end{equation}

@@

However, the sparsity of $\Delta\theta$, see \eqref{eq:nparam:lrm:bias:sparse}, can be exploited such that the bias on the \gls{FRF} can be expressed as:
\begin{equation}
  \Delta \LocalModel{G}(\theta,\dW)
  \approx
  \frac{\LocalModel{G}(\theta, \dW)}
           {\LocalModel{A}(\theta, \dW)}
  \begin{bmatrix}
    \dW^1 & \deemph{\cdots} & \dW^{\order{A}}
  \end{bmatrix}
  \Delta \theta_A
  \text{.}
\end{equation}
Equivalently, the relative bias of the \gls{FRF} can be written as
\begin{equation}
\frac{\Delta \LocalModel{G}(\theta,\dW)}{\LocalModel{G}(\theta,\dW)}
\approx
\LocalModel{A}(\theta,\dW)^{-1}
\begin{bmatrix}
    \dW^1 & \deemph{\cdots} & \dW^{\order{A}}
  \end{bmatrix}
  \left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}
  \begin{bmatrix}
  \Delta\theta_A\\
  0\\
  0\\
  \end{bmatrix}
  \text{.}
\end{equation}

@@

\begin{table}
\centering
\caption{Local modeling methods used in the Monte Carlo analysis.}
\label{tbl:nparam:methods}
\begin{tabular}{lccccr} 
\toprule
\textsc{Method} & \textsc{Color} & $\nWind$ & $\nth$ & \DOF & $\Delta\SNR \axisunit{dB}$ \\
\midrule
 $\ETFE$ &\color{etfe} --- & 1  & 1  & 0 & 0 \\
 $\lrm{7,1,1,1}$ & \color{lrm7111}\bfseries --- & 15  & 5 & 10  & 4.77 \\
 $\lric{7,2,2,2}$ &\color{lric7222} \bfseries --- &   15 & 8 & 7  & 2.73 \\
 $\lrm{7,2,2,2}$ &\color{lrm7222} \bfseries --- &   15 & 8 & 7  & 2.73 \\
 $\lrm{6,2,2,2}$ &\color{lrm6222} \bfseries --- &   13 & 8 & 5  & 2.11 \\
 $\lpm{7,4,4}$ &\color{lpm744} \bfseries --- &   15 & 10 & 5  & 1.76 \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{guideline}[Use the LRM for resonant systems when $\SNR > 10 \unit{dB}$]
For resonant systems, the \gls{LRM} offers better performance than \gls{LPM} when a good \gls{SNR} is obtained in the frequency domain.
When the \gls{SNR} is poor ($\leq 10\unit{dB}$), the \gls{LPM} may perform a bit better than \gls{LRM}.
\end{guideline}

\begin{guideline}[For long measurements, LPM and LRM are similar]
For very long measurements, the difference between \gls{LPM} and \gls{LRM} becomes negligible if the different resonances are well-separated.
\end{guideline}

@@

  % \section{Measurements}
  % \label{sec:measurements}
  %  \subsection{AVIS}
   % \subsection{F16}

### 2016-04-23 ch03-lrm/nonparametric.tex
Clean up most definitions
Actually include MC figures
Some things
Stuff :-)

\chapter{Non-parametric Modeling}
\def\thisDir{ch03-lrm}
\tikzsetfigurename{ch03fig}
\label{sec:nonparametric}
% \myEpigraph{}{}{}

@@

Whereas many areas of system identification focus on parametric identification and/or parameter estimation, non-parametric techniques have recently received new research interest from the community.

@@

\input{\thisDir/proofLOOCV.tex}

>>>>

\chapter{Non-parametric Modeling}
\def\thisDir{ch03-lrm}
\tikzsetfigurename{ch03fig}
\label{sec:nonparametric}
\myEpigraph{I love deadlines. I like the whooshing sound they make as they fly by.
}{Douglas Adams}{}

@@

Whereas many areas of system identification focus on parametric identification and/or parameter estimation, non-parametric techniques have recently received renewed interest from the community.

@@

% \input{\thisDir/proofLOOCV.tex}
